### 2017-06-15 Manuscript/Response_to_reviewers.Rmd

\bigskip

**L 247: Please compare cross-sectional and longitudinal telomere shortening and discuss it.**

RESPONSE: We now include a formal test of the within and between individual slopes, as described in van de Pol and Wright (2009), and find that there is no difference between cross-sectional and longitudinal telomere shortening - lines XX:

\textit{"Further, we found no evidence that within and between individual slopes of telomere shortening varied (estimate = -0.003, CIs = -0.008, 0.003; see Van de Pol and Wright 2009), suggesting that there was no difference between cross-sectional and longitudinal telomere shortening with age."}

@@

---

\bigskip

**L 340: Does telomere lengthening or shortening relate to environmental conditions in your study?**

RESPONSE: As stated in the discussion (lines XX), we feel that this is an important area for future research, and we hope that it will be a whole manuscript on its own. We are already pushing the page limit for Journal of Animal Ecology, so adding in an entire new set of analyses is just not possible.

### 2017-06-15 Manuscript/SW_early_life_telomeres_FULL.Rmd
Edits

A within-individual analysis of RTL and age revealed that the top model explaining RTL contained $\Delta$log age, which reflects within-individual changes in log-transformed age (Table 1B). Models including cohort ID were substantially poorer fits than a model only containing age (Table 1B). RTL decreased with $\Delta$log age (estimate = `r round3(fixef(RTL_DeltaLogAge)['DeltaLogAge'],lessthan = F)`, CIs = `r paste(round3(CIwald(RTL_DeltaLogAge)['DeltaLogAge',],lessthan = F),collapse = ', ')`), confirming that within-individual telomere shortening occurs across the Seychelles warbler dataset. Further, we found no evidence that within and between individual slopes of telomere shortening varied (``r round3(fixef(RTL_LogAge2)['MeanAge'],lessthan = F)`, CIs = `r paste(round3(CIwald(RTL_LogAge2)['MeanAge',],lessthan = F),collapse = ', ')`; see Van de Pol and Wright 2009)

>>>>

A within-individual analysis of RTL and age revealed that the top model explaining RTL contained $\Delta$log age, which reflects within-individual changes in log-transformed age (Table 1B). Models including cohort ID were substantially poorer fits than a model only containing age (Table 1B). RTL decreased with $\Delta$log age (estimate = `r round3(fixef(RTL_DeltaLogAge)['DeltaLogAge'],lessthan = F)`, CIs = `r paste(round3(CIwald(RTL_DeltaLogAge)['DeltaLogAge',],lessthan = F),collapse = ', ')`), confirming that within-individual telomere shortening occurs across the Seychelles warbler dataset. Further, we found no evidence that within and between individual slopes of telomere shortening varied (estimate = `r round3(fixef(RTL_LogAge2)['MeanAge'],lessthan = F)`, CIs = `r paste(round3(CIwald(RTL_LogAge2)['MeanAge',],lessthan = F),collapse = ', ')`; see Van de Pol and Wright 2009), suggesting that there was no difference between cross-sectional and longitudinal telomere shortening with age.
